---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rlowdb

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/feddelegrand7/rlowdb/branch/main/graph/badge.svg)](https://app.codecov.io/gh/feddelegrand7/rlowdb?branch=main)
[![R-CMD-check](https://github.com/feddelegrand7/rlowdb/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/feddelegrand7/rlowdb/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`rlowdb` is a lightweight, JSON-based database for R, inspired by [LowDB](https://github.com/typicode/lowdb) from JavaScript. It provides a simple and efficient way to store, retrieve, update, and delete structured data without the need for a full database system.

## Features

- **Lightweight & File-Based**: Uses JSON for persistent storage.
- **Easy-to-Use API**: Supports CRUD operations (Create, Read, Update, Delete).
- **Flexible Queries**: Allows filtering with expressive conditions.
- **No External Dependencies**: No need for SQL or additional database software.

## Installation

`rlowdb` is not yet available on CRAN, but you can install it from GitHub:

```{r, eval=FALSE}
devtools::install_github("feddelegrand7/rlowdb")
```


## Usage
### Initializing the Database
To start using `rlowdb``, create a new database instance by specifying a JSON file:

```{r}
library(rlowdb)
db <- rlowdb$new("DB.json")
```

### Inserting Data

The `insert` method takes two parameters, a `collection` and a `record`, think of the `collection` parameter as a `table` in the __SQL__ world. Think of the `record` parameter as a `list` of names, each name/value pair representing a specific column and it's value.

Add records to a collection:

```{r}
db$insert(
  collection = "users", 
  record = list(id = 1, name = "Alice", age = 30)
)
db$insert(
  collection = "users", 
  record = list(id = 2, name = "Bob", age = 25)
)

db$insert(
  collection = "users", 
  record = list(id = 3, name = "Alice", age = 30)
)
```


### Retrieving Data

Get all stored data:

```{r}
db$get_data()
```

Find a specific record:

```{r}
db$find(collection = "users", key = "id", value = 1)
```

### Updating Records

Modify existing records:

```{r}
db$update(
  collection = "users", 
  key = "id", 
  value = 1, 
  new_data = list(age = 31)
)

db$get_data()
```

The `upsert` methods allows you to update a record if it exists, otherwise, it will be inserted:

```{r}
db$upsert(
  collection = "users", 
  key = "id", 
  value = 1, 
  new_data = list(age = 25)
)

db$get_data()
```

```{r}
db$upsert(
  collection = "users", 
  key = "id", 
  value = 100, 
  new_data = list(age = 25)
)

db$get_data()
```


### Deleting Records

```{r}
db$delete(collection = "users", key = "id", value = 100) 

db$get_data()
```

### Querying Data

Find users older than 25:

```{r}
db$query(collection = "users", condition = "age > 25")
```

Query with multiple conditions:

```{r}
db$query(collection = "users", condition = "age > 20 & id > 1")
```


### Listing the collections

The `list_collections` method returns the names of the collections within your DB:

```{r}
db$list_collections()
```

### Counting 

Using the `count` method, you can get the number of records a collection has:

```{r}
db$count(collection = "users") 
```


### Check if exists

It possible to verify if a `collection`, a `key` or a `value` exists within your `DB`:

```{r}
db$exists_collection(collection = "users")
```

```{r}
db$exists_collection(collection = "nonexistant")
```
```{r}
db$exists_key(collection = "users", key = "name")
```
```{r}
db$exists_value(
  collection = "users",
  key = "name",
  value = "Alice"
)
```

```{r}
db$exists_value(
  collection = "users",
  key = "name",
  value = "nonexistant"
)
```

### Clear, Drop Data

It is possible to `clear` a collection. This will remove all the elements belonging to the collection but not drop the collection it self: 

```{r}
db$insert(collection = "countries", record = list(id = 1, country = "Algeria", continent = "Africa"))

db$insert(collection = "countries", record = list(id = 1, country = "Germany", continent = "Europe"))

db$get_data()
```

Now, look what happened when we use the `clear` method on the `countries` collection: 

```{r}
db$clear("countries")

db$get_data()
```


Using the `drop` method, one can drop a whole collection: 

```{r}
db$drop(collection = "countries")
db$get_data()
```

Finally, `drop_all` will drop all the `collections` within your `DB`: 


```{r}
db$drop_all()
db$get_data()
```

### Error Handling

`rlowdb` provides error handling for common issues. For example, attempting to update a collection that does not exist will result in an informative error: 

```{r, error=TRUE}
db$update(
  collection = "nonexistant", 
  key = "id",
  value = 1, 
  new_data = list(age = 40)
)  
```

### Future Features

- Support for nested data structures.
- More advanced query capabilities.
- Compatibility with alternative file formats (e.g., CSV, SQLite).


## Code of Conduct

Please note that the ralger project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.



```{r, echo=FALSE}
unlink("DB.json")
```






